Software Citation Principles

### 2016-03-29 sc-principles/software-citation-principles.tex
edits in a few scattered places - removing some old comments

These software citation principles were originally based on an adaptation of the FORCE11 Data Citation Principles~\cite{data-citation-principles}, and then were modified based on working group discussions and the information from the use cases and the related work.
The adaptations and modifications have been made because software, while similar to data in terms of not traditionally having been cited in publications, is also different than data in that it can be used to express or explain concepts, it is updated more frequently, and it is executable.  \katznote{are there any other difference between software and data that should be stated here?}  Also, while software can be considered a type of data, the converse is not generally true.

@@

Perhaps in recognition of the broad range of research domains struggling with
the challenge of better recognizing the role of software, a number of community
efforts hosted (and sponsored) by funders and agencies in both the US (e.g., NSF,
NIH, Alfred P.\ Sloan Foundation) and UK (e.g., SFTC, JISC) have run a number of
workshops with participants from across a range of disciplines.

Most notable of the community efforts are those of WSSSPE~\cite{wssspe} and
SSI~\cite{ssi-workshops} who between them have run a series of workshops aimed
at gathering together community members with an interest in improving the
status of software in academia. \asnote{Not sure this is a sufficient one-line
summary but couldn't think of anything better.} In each of the three years that
WSSSPE workshops have run thus far, the participants have produced a report
documenting the topics covered. Section 5.8 and Appendix J in \cite{WSSSPE3}
has some preliminary work and discussion particularly relevant to this working
group.
Up
In addition, a number of academic publishers such as APA~\cite{APA-guidelines} have recommendations for submitting authors on how to cite software and journals such as F1000Research~\cite{F1000}, SoftwareX~\cite{softwareX}, and Open Research Computation~\cite{softwareX} allow for submissions entirely focussed on research software.

@@

The Life sciences: In May 2014, the NIH held a workshop aimed at helping the biomedical community discover, cite and reuse software written by their peers. The primary outcome of this workshop was the Software Discovery Index Meeting Report~\cite{software-discovery-index}. The authors of the report cite discuss what framework would be required for supporting a Software Discovery Index including the need for unique identifiers, how citations to these would be handled by publishers and the critical need for metadata to describe software packages.

@@

Kyle E.~Niemeyer, Oregon State University

>>>>

These software citation principles were originally based on an adaptation of the FORCE11 Data Citation Principles~\cite{data-citation-principles}, and then were modified based on working group discussions and the information from the use cases and the related work.
The adaptations and modifications have been made because software, while similar to data in terms of not traditionally having been cited in publications, is also different than data in that it can be used to express or explain concepts, it is updated more frequently, and it is executable. 
% \katznote{are there any other difference between software and data that should be stated here?}  
Also, while software can be considered a type of data, the converse is not generally true.

@@

Perhaps in recognition of the broad range of research domains struggling with
the challenge of better recognizing the role of software, a number of community
efforts hosted (and sponsored) by funders and agencies in both the US (e.g., NSF,
NIH, Alfred P.\ Sloan Foundation) and UK (e.g., SFTC, JISC, Wellcome Trust) have run a number of
workshops with participants from across a range of disciplines.

Most notable of the community efforts are those of WSSSPE~\cite{wssspe} and
SSI~\cite{ssi-workshops}, who between them have run a series of workshops aimed
at gathering together community members with an interest in 1) defining the set of problems
related to the role of software and associated people in research settings, particularly academia,
2) discussing potential solutions to those problems, and 3) beginning to work on implementing
some of those solutions.
%\asnote{Not sure this is a sufficient one-line
%summary but couldn't think of anything better.} \katznote{I've edited it a bit}
In each of the three years that
WSSSPE workshops have run thus far, the participants have produced a report
documenting the topics covered. Section 5.8 and Appendix J in \cite{WSSSPE3}
has some preliminary work and discussion particularly relevant to this working
group.
Up
In addition, a number of academic publishers such as APA~\cite{APA-guidelines} have recommendations for submitting authors on how to cite software and journals such as F1000Research~\cite{F1000}, SoftwareX~\cite{softwareX}, and Open Research Computation~\cite{softwareX} allow for submissions entirely focussed on research software.

@@

The life sciences: In May 2014, the NIH held a workshop aimed at helping the biomedical community discover, cite and reuse software written by their peers. The primary outcome of this workshop was the Software Discovery Index Meeting Report~\cite{software-discovery-index}. The authors of the report cite discuss what framework would be required for supporting a Software Discovery Index including the need for unique identifiers, how citations to these would be handled by publishers and the critical need for metadata to describe software packages.

@@

Kyle E.~Niemeyer, Oregon State University (co-chair)
